---
title: "Functions"
author: "Lorena Garcia-Perez (lorena.garcia-perez@crick.ac.uk)"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
---

## Custom functions for plotting

### 1. Themes

Chose white or dark themes as part of the plotting funtion inputs  

```{r}


basicTheme <- list(
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, margin = margin(t = 0, r = 0, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    strip.text.x = element_text(colour = "black", size = 12),
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.title = element_text(size = 10),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent", colour = "transparent"),
    strip.background = element_rect(fill = "transparent", colour = "transparent"),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(size = 0.25, color = "gray85"),
    panel.grid.major.y = element_line(size = 0.25, color = "gray85"),
    axis.ticks = element_blank()
  )
)

whiteTheme <- list(theme())

darkTheme <- list(
  theme(
    axis.text = element_text(colour = "white"),
    axis.title = element_text(colour = "white"),
    strip.text.x = element_text(colour = "white"),
    panel.border = element_rect(color = "white")
  )
)
```


### 2. Box plots: RNA copy number and RNA concentration

```{r}
box.plots <-
  function(data, point.color = "gray40", Theme = whiteTheme) {
    data$day <- as.factor(data$day)
    levels(data$day) <- days.names

    p.copy <-
      ggplot(data, aes(day, copy, group = day)) +
      geom_quasirandom(
        shape = 20, color = point.color,
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_boxplot(aes(color = channel),
        outlier.shape = NA, fill = NA, lwd = 0.5
      ) +
      scale_color_manual(values = col.custom) +
      scale_y_continuous(
        expand = c(0, 0),
        breaks = seq(0, 150, 25),
        limits = c(
          0, # max(data$copy)
          150
        )
      ) +
      labs( # subtitle = "RNA copy number",
        x = "",
        y = "RNAs / cell"
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      theme(
        legend.position = "none",
        panel.grid.major.x = element_blank()
      ) +
      # theme(plot.subtitle=element_text(colour = colourPalette[1])) +
      facet_wrap(~channel)
    # print(p.copy)

    p.conc <-
      ggplot(data, aes(day, density.volume, group = day)) +
      geom_quasirandom(
        shape = 20, colour = point.color,
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_boxplot(aes(color = channel),
        outlier.shape = NA, fill = NA, lwd = 0.5
      ) +
      scale_color_manual(
        values = col.custom,
        labels = c(" ", " ", " ")
      ) +
      scale_y_continuous(
        expand = c(0, 0),
        breaks = seq(0, max(data$density.volume) + 0.01, 0.02),
        limits = c(0, max(data$density.volume) + 0.01)
      ) +
      labs( # subtitle = "RNA concentration",
        x = "",
        y = bquote("RNAs / " ~ mu * m^3)
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      theme(
        legend.position = "none",
        strip.text.x = element_blank(),
        panel.grid.major.x = element_blank()
      ) +
      # theme(plot.subtitle=element_text(colour = colourPalette[5])) +
      facet_wrap(~channel)

    panel <- grid.arrange(p.copy, p.conc, nrow = 2, heights = c(1.15, 1))
    ggsave(paste("./figures/copy_concentration_box_plots_",
      deparse(substitute(Theme)), ".pdf",
      sep = ""
    ),
    panel,
    width = 9, height = 6
    )
  }

box.plot.area.volume <-
  function(data, point.color = "gray40", Theme = whiteTheme) {
    data$day <- as.factor(data$day)
    levels(data$day) <- days.names

    p.cell.area <-
      ggplot(filter(data, channel == genes[1]), aes(day, cell.area, group = day)) +
      geom_quasirandom(
        shape = 20, color = point.color,
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_boxplot(aes(color = channel),
        outlier.shape = NA, fill = NA, lwd = 0.5
      ) +
      scale_color_manual(values = "tan1") +
      scale_y_continuous(
        expand = c(0, 0),
        breaks = seq(0, max(data$cell.area), 100),
        limits = c(0, max(data$cell.area))
      ) +
      labs( # subtitle = "Cell area"",
        x = "",
        y = bquote("Cell area [" ~ mu * m^2 ~ "]")
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      theme(
        legend.position = "none",
        strip.text.x = element_blank(),
        panel.grid.major.x = element_blank()
      )

    p.cell.volume <-
      ggplot(filter(data, channel == genes[1]), aes(day, cell.volume, group = day)) +
      geom_quasirandom(
        shape = 20, color = point.color,
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_boxplot(aes(color = channel),
        outlier.shape = NA, fill = NA, lwd = 0.5
      ) +
      scale_color_manual(values = "tan2") +
      scale_y_continuous(
        expand = c(0, 0),
        breaks = seq(0, max(data$cell.volume), 500),
        limits = c(0, max(data$cell.volume))
      ) +
      labs( # subtitle = "Cell volume"",
        x = "",
        y = bquote("Cell volume [" ~ mu * m^3 ~ "]")
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      theme(
        legend.position = "none",
        strip.text.x = element_blank(),
        panel.grid.major.x = element_blank()
      )

    panel <- grid.arrange(p.cell.area, p.cell.volume, ncol = 2)
    ggsave(paste("./figures/area_volume_box_plot_",
      deparse(substitute(Theme)), ".pdf",
      sep = ""
    ),
    panel,
    width = 6, height = 3
    )
  }
```

### 3. Scatter plots: RNA copy number and RNA concentration

channel.1 and channel.2 must be supplied as strings  
data must have channel as factor  

```{r}
pairwise.scatter.plot <-
  # Could try to make axis limit specific for pair of genes being plot each time
  function(data, channel.1, channel.2,
             channel.color, Theme = whiteTheme) {
    data.formatted <-
      data %>%
      mutate(day = factor(day)) %>%
      mutate(day = mapvalues(day, from = days, to = days.names))

    data.copy <-
      data.formatted %>%
      select(channel, day, image, cell.id, cell.area, copy) %>%
      spread(channel, copy)

    p.copy <-
      ggplot(data.copy, aes_string(channel.2, channel.1)) +
      geom_point(
        shape = 20, aes_string(color = channel.color),
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_density_2d(size = 0.25, bins = 15, color = "tan1") +
      scale_color_viridis(
        option = "inferno",
        name = bquote(atop(.(channel.color), "[ RNAs / cell ]"))
      ) +
      labs( # title = "RNA copy number",
        x = bquote(atop(.(channel.2) ~ " [ RNAs / cell ]", " ")),
        y = bquote(.(channel.1) ~ " [ RNAs / cell ]")
      ) +
      scale_x_continuous(
        # breaks = seq(0, max(data$copy), 25),
        # limits = c(0, max(data$copy)),
        expand = c(0, 0)
      ) +
      scale_y_continuous(
        # breaks = seq(0, max(data$copy), 25),
        # limits = c(0, max(data$copy),
        expand = c(0, 0)
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      # theme(plot.title = element_text(colour = colourPalette[1])) +
      facet_wrap(~day)

    data.conc <-
      data.formatted %>%
      select(channel, day, image, cell.id, cell.area, density.volume) %>%
      spread(channel, density.volume)

    p.conc <-
      ggplot(data.conc, aes_string(channel.2, channel.1)) +
      geom_point(
        shape = 20, aes_string(colour = channel.color),
        size = 1.25, alpha = 0.7, stroke = 0
      ) +
      geom_density_2d(size = 0.25, bins = 15, color = "tan1") +
      scale_color_viridis(
        option = "inferno",
        name = bquote(atop(.(channel.color), "[ RNAs /" ~ mu * m^3 ~ "]"))
      ) +
      labs( # title = "RNA concentration",
        x = bquote(.(channel.2) ~ " [ RNAs /" ~ mu * m^3 ~ "]"),
        y = bquote(.(channel.1) ~ " [ RNAs /" ~ mu * m^3 ~ "]")
      ) +
      scale_x_continuous(
        # breaks = seq(0, max(data$density), 1),
        # limits = c(0, max(data$density)),
        expand = c(0, 0)
      ) +
      scale_y_continuous(
        # breaks = seq(0, max(data$density), 1),
        # limits = c(0, max(data$density)),
        expand = c(0, 0)
      ) +
      theme_linedraw(base_size = 12) +
      basicTheme +
      Theme +
      theme(strip.text.x = element_blank()) +
      # theme(plot.title = element_text(colour = colourPalette[5])) +
      facet_wrap(~day)

    panel <- grid.arrange(p.copy, p.conc, nrow = 2, heights = c(1.25, 1))
    ggsave(
      paste("./figures/copy_concentration_pairwise_scatter_plots_",
        deparse(substitute(Theme)), "_", channel.2, "_", channel.1, ".pdf",
        sep = ""
      ),
      panel,
      width = 9, height = 6
    )
  }
```




## Custom functions to estimate RNA half-life

Trancription probability  

```{r}
prob.TX.fn <- function(data, day.chosen, gene.chosen, ploidy) {
  # day.chosen and gene.schosen must be input as character (for example, "4"
  # and "Olig2")
  # ploidy must be input as numeric
  day.gene.TX.subset <-
    pull(filter(data, day == day.chosen & channel == gene.chosen), TX.number)
  return(sum(day.gene.TX.subset) / (ploidy * length(day.gene.TX.subset)))
}
```

Mean RNA copy number  

```{r}
mean.copy.fn <- function(data, day.chosen, gene.chosen) {
  # day.chosen and gene.schosen must be input as character (for example, "4"
  # and "Olig2")
  day.gene.copy.subset <-
    pull(filter(data, day == day.chosen & channel == gene.chosen), copy)
  return(mean(day.gene.copy.subset))
}
```

Mean number of RNA pol = mean number of molecules at TX spot  

```{r}
mean.nPol.fn <- function(data, day.chosen, gene.chosen) {
  # day.chosen and gene.schosen must be input as character (for example, "4"
  # and "Olig2")
  day.gene.nPol.subset <-
    pull(filter(data, day == day.chosen & channel == gene.chosen), TX.mol)
  return(mean(day.gene.nPol.subset))
}
```


