---
title: "smFISH analysis"
author: "Lorena Garcia-Perez (lorena.garcia-perez@crick.ac.uk)"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
geometry: margin = 2 cm

output:
  
  pdf_document:
    latex_engine: xelatex
    highlight: haddock
    toc: true
    
#“default”, “tango”, “pygments”, “kate”, “monochrome”, “espresso”, “zenburn”, “haddock”

knit: (function(inputFile, encoding) { 
      out_dir <- "docs";
      rmarkdown::render(inputFile,
                        encoding = encoding, 
                        output_file = file.path(dirname(inputFile),
                                      out_dir, "smFISH_analysis.pdf")) })
---

# Set up:

## Load libraries  
  
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(ggbeeswarm)
require(plyr) # I use it for mapValues in pairwise.scatter custom function
library(tidyverse)
# library(tibble) #for add_column etc
library(knitr)
library(rmarkdown)
library(pander)
library(jsonlite)
library(gridExtra)
library(LaCroixColoR)

library(viridis)

#Rstudio addins > style active file (uses the styler package)
```

## Global output doc settings

```{r}
opts_chunk$set(
  panderOptions("table.split.table", 120),
  fig.align = "center",
  fig.width = 7
)
# options(dplyr.print_min = 6, dplyr.print_max = 6)
```

# Import file that loads data and contains user-defined parameters
  
```{r child = "load_clean.Rmd"}
render("load_clean.Rmd")
```

# Import file with custom functions
  
```{r child = "func.Rmd"}
render("func.Rmd")
```

# RNA copies, concentration and transcription per cell

## RNA copy number and concentration

```{r}

# detach("package:plyr", unload=TRUE)

copy.density.cell.nonZero <-
  file %>%
  mutate(cell.area = cell.area * pixel.size) %>% # Convert cell area in pixels to cell area in µm^2
  filter(spot.id != 0) %>%
  group_by(channel, day, image, cell.id, cell.area) %>%
  summarise(copy = n()) %>% # RNA copy number per cell
  mutate(density = round(copy / cell.area, 3)) # RNA densities in terms of RNA molecules per µm^2

copy.density.cell.Zero <-
  file %>%
  mutate(cell.area = cell.area * pixel.size) %>% # Convert cell area in pixels to cell area in µm^2
  filter(spot.id == 0) %>%
  group_by(channel, day, image, cell.id, cell.area) %>%
  summarise(copy = 0) %>% # RNA copy number per cell
  mutate(density = round(copy / cell.area, 3)) # RNA densities in terms of RNA molecules per µm^2

copy.density.cell <- full_join(copy.density.cell.nonZero, copy.density.cell.Zero)

copy.density.cell <-
  copy.density.cell %>%
  mutate(cell.volume = cell.area * Z.dimensions) %>%
  mutate(density.volume = copy / cell.volume)

pander(head(copy.density.cell), style = "rmarkdown")
```
  
## Transcriptional parameters

To determine which spots can be considered transcription spots, we need to know 
the area and maximum intensity values for all the spots. Then we normalize
according to each channel, day and image  
  
```{r}
TX.threholds <-
  file %>%
  group_by(channel, day, image) %>%
  # Choose threholds specifically for each image and channel
  summarise(
    MaxI.median = median(max.int), MaxI.mad = mad(max.int),
    area.median = median(area), area.mad = mad(area)
  ) %>%
  mutate(
    MaxI.thres = MaxI.median + 3 * MaxI.alpha * MaxI.mad,
    area.thres = area.median + 3 * area.alpha * area.mad
  )
pander(head(TX.threholds), style = "rmarkdown")
```
  
```{r}
TX.number.cell <-
  file %>%
  group_by(channel, day, image) %>%
  left_join(TX.threholds) %>%
  mutate(is.TX.spot = max.int > MaxI.thres & area > area.thres) %>%
  group_by(channel, day, image, cell.id) %>%
  summarise(TX.number = sum(is.TX.spot))
pander(head(TX.number.cell), style = "rmarkdown")
```

To estimate how many RNA molecules are within each TX spot, we take into account
the mean intensity and area of each TX spot, and divide by the median values of
all the spots in the same channel, day and image. A correction factor that takes
into account probe location along the RNA is also applied.

```{r}
TX.mol <-
  file %>%
  group_by(channel, day, image) %>%
  left_join(TX.threholds) %>%
  mutate(is.TX.spot = max.int > MaxI.thres & area > area.thres) %>%
  mutate(meanInt.median = median(mean.int)) %>%
  filter(is.TX.spot == TRUE) %>%
  mutate(probe.loc = channel) %>%
  mutate(probe.loc = as.numeric(recode(probe.loc,
    "0" = correction.factor[1],
    "1" = correction.factor[2],
    "2" = correction.factor[3]
  ))) %>%
  mutate(TX.mol = mean.int * area / (probe.loc * meanInt.median * area.median))
TX.mol$channel <- as.factor(TX.mol$channel)
levels(TX.mol$channel) <- genes
pander(head(TX.mol), style = "rmarkdown")
```
  
# Joining datasets and subsetting cell population

## All cells

```{r}
single.cell.data <-
  inner_join(copy.density.cell, TX.number.cell) %>%
  ungroup()
single.cell.data$channel <- as.factor(single.cell.data$channel)
levels(single.cell.data$channel) <- genes

setwd(parent)
write.csv(single.cell.data, "./ouput/single_cell_data.csv", row.names = FALSE)
setwd(wd)

pander(head(single.cell.data), style = "rmarkdown")
```

```{r}
single.cell.data.nested <-
  single.cell.data %>%
  group_by(day, image, cell.id, cell.area, channel) %>%
  nest() %>%
  group_by(day, image, cell.id, cell.area) %>%
  nest() %>%
  group_by(day) %>%
  nest()

myjson <- toJSON(single.cell.data.nested, pretty = TRUE, auto_unbox = TRUE)

setwd(parent)
write(myjson, "./ouput/single_cell_data_with_TX.json")
setwd(wd)

# json_data <- fromJSON(file = "./ouput/single_cell_data_with_TX.json")
```

```{r}
# This is the more complete, non subseted, dataset
# But, if a cell has two TX spots, it will appear on two rows and so on
withTXmol.single.cell.data <-
  TX.mol %>%
  select(channel, day, image, cell.id, spot.id, TX.mol) %>%
  full_join(single.cell.data) %>%
  rename(spot.id.TX = spot.id) %>%
  arrange(channel, day, image) %>%
  replace_na(list(TX.mol = 0))
pander(head(withTXmol.single.cell.data), style = "rmarkdown")
```

```{r}
two.alleles.data <-
  TX.mol %>%
  select(channel, day, image, cell.id, TX.mol) %>%
  group_by(channel, day, image, cell.id) %>%
  mutate(freq = n()) %>%
  ungroup() %>%
  filter(freq == 2) %>%
  select(-freq) %>%
  mutate(TX.id = rep(c(1, 2), length.out = length(.$channel))) %>%
  # Create a column with appropiate number of rows
  spread(TX.id, TX.mol) %>%
  rename(TX.id.1 = "1", TX.id.2 = "2")
two.alleles.data$channel <- as.factor(two.alleles.data$channel)
levels(two.alleles.data$channel) <- genes
pander(head(two.alleles.data), style = "rmarkdown")
```
  
## Cells expressing certain genes


single.cell.expressing.data <-
  single.cell.data %>%
  filter(density >= 1)
pander(head(single.cell.expressing.data), style = "rmarkdown")



single.cell.sox2.expressing.data <-
  # This could be combined within single.cell.progenitors.data with filter_if()?
  single.cell.data %>%
    filter(channel == progenitors.gene) %>%
    filter(density >= 1) # Require sox2 density >= 1 molecule / µm^2
pander(head(single.cell.sox2.expressing.data), style = "rmarkdown")


## Neural progenitor cells

This dataset will be used for posterior analysis


single.cell.progenitors.data <-
  # Select progenitors according to sox2 density >= 1 molecule / µm^2
  semi_join(single.cell.data, single.cell.sox2.expressing.data,
    by = c("day", "image", "cell.id", "cell.area")
  )
pander(head(single.cell.progenitors.data), style = "rmarkdown")

  


# Plotting RNA copy number and concentration for progenitor cells

```{r}
genes.pairwise <-
  tibble(genes) %>%
  rename(gene.1 = genes) %>%
  mutate(gene.2 = gene.1) %>%
  expand(gene.1, gene.2) %>%
  filter(gene.1 < gene.2) %>% # To get unique combinations
  mutate(gene.3 = c(genes[2], genes[3], genes[1]))
```

```{r}
# opts_knit$set(root.dir = parent)
setwd(parent)

box.plot.area.volume(
  single.cell.data
  # ,"white",darkTheme
)

# copy and concentration box plots
box.plots(
  single.cell.data
  # ,"white",darkTheme
)
```
  
```{r}

setwd(parent)

for (g in c(1, 2, 3)) {
  pairwise.scatter.plot(
    single.cell.data,
    genes.pairwise$gene.1[g], genes.pairwise$gene.2[g],
    genes.pairwise$gene.3[g]
    # , darkTheme
  )
}
```



# Calculating RNA half-life based on smFISH data

single.cell.progenitors.data is used, but results with single.cell.data are
quite similar

```{r}
mean.properties.data <-
  tibble(
    day = character(), gene = character(), prob.TX = as.numeric(),
    mean.copy = as.numeric(), mean.nPol = as.numeric()
  )

for (d in days) {
  for (g in genes) {
    mean.properties.data <-
      add_row(mean.properties.data,
        day = d, gene = g,
        prob.TX =
          round(prob.TX.fn(single.cell.data, d, g, ploidy), 3),
        mean.copy =
          round(mean.copy.fn(single.cell.data, d, g), 3),
        mean.nPol =
          round(mean.nPol.fn(TX.mol, d, g), 3)
      )
  }
}
mean.properties.data <- mean.properties.data %>% replace_na(list(mean.nPol = 0))

mean.properties.data <-
  mean.properties.data %>%
  mutate(
    vPol = vPol.chosen, gene.length = rep(genes.length, 3),
    ploidy = ploidy
  ) %>%
  mutate(TX.rate = mean.nPol * vPol / gene.length) %>%
  mutate(deg.rate = prob.TX * TX.rate * ploidy / mean.copy) %>%
  mutate(half.life = log(2) / deg.rate) %>% # in min
  arrange(gene, day)

pander(mean.properties.data, style = "rmarkdown")
```
  
  
# Effect of noise  
  
```{r}
prob.TX.pairs.data <-
  mean.properties.data %>%
  select(day, gene, prob.TX) %>%
  rename(gene.1 = gene, prob.TX.1 = prob.TX) %>%
  mutate(gene.2 = gene.1) %>%
  expand(nesting(day, gene.1, prob.TX.1), gene.2) %>%
  group_by(day) %>%
  mutate(prob.TX.2 = rep(unique(prob.TX.1), times = 3)) %>%
  filter(gene.1 < gene.2) # To get unique combinations
pander(head(prob.TX.pairs.data), style = "rmarkdown")
```
  


TO BE FINISHED...  

for (d in days) {
  for (g in genes) {
    prob.TX.pairs.data$prob.TX.both <-
      add_row(prob.TX.pairs.data$prob.TX.both,
              prob.TX.pairs.fn())
              
  
```{r}
prob.TX.pairs.fn <-
  function(data, day.chosen, gene.chosen.1, gene.chosen.2, ploidy) {
    # day.chosen and gene.schosen must be input as character (for example, "4"
    # and "Olig2")
    # ploidy must be input as numeric
    day.gene.TX.subset <-
      filter(data, day == day.chosen &
        (channel == gene.chosen.1 | channel == gene.chosen.2))
    dual.TX <-
      day.gene.TX.subset %>%
      select(channel, image, day, cell.id, TX.number) %>%
      filter(TX.number != 0) %>%
      group_by(day, image, cell.id) %>%
      mutate(freq = n()) %>%
      ungroup() %>%
      filter(freq == 2) %>%
      select(-freq) %>%
      spread(channel, TX.number) %>%
      return(nrow(dual.TX) / (nrow(day.gene.TX.subset)))
    # Here I do not make use of the ploidy concept...
  }
```
  
two.alleles.data$channel <- as.factor(two.alleles.data$channel )
levels(two.alleles.data$channel) <- genes
  
TX.probability.all[,6] <- TX.probability.all$TX.prob.pair /
  (TX.probability.all$TX.prob.oneC * TX.probability.all$TX.prob.otherC)
colnames(TX.probability.all)[6] <- c("extrinsic.noise")
  
To measure the influence of extrinsic noise:  
THIS SHOULD BE DONE WITH A CONSTITUTIVE GENE TO COMPARE!!!  
  
###########################################################################  

source("26_cell_area.R")  
  
###########################################################################
###########################################################################
###########################################################################
###########################################################################
  
file$channel <- as.factor(file$channel)
levels(file$channel) <- genes

rect <- data.frame(xmin=threshold.area, xmax=Inf,
                   ymin=threshold.MaxInt, ymax=Inf,
                   channel=c(0,1,2))
rect$channel <- as.factor(rect$channel)
levels(rect$channel) <- genes

###########################################################################
###########################################################################

source("26_spot_area_maxInt.R")


source("26_TXnumber_vs.R")


source("26_barTXprob.R")


source("26_barTXprobExtrinsic.R")


source("26_boxTXmolecules.R")


source("26_TX2alleles.R")

###########################################################################
###########################################################################

plots <- c(hist.spots.area,
           hist.spots.MaxInt,
           p.spots.AreaVsMaxInt,
           p.TX.numberVsCopy,
           p.TX.numberVsDensity,
           p.TX.numberVsCellArea,
           bar.TX.prob,
           bar.TX.prob.extrinsic,
           box.TX.molecules,
           p.TX.2alleles.molecules.final)

###########################################################################
###########################################################################

Heterogeneity analysis

var(spots.cell$copy.C1) / mean (spots.cell$copy.C1)

#Fano factor = variance / mean = sigma^2 / mean
FF.expressing <- data.frame(FF.C0=as.numeric(),
                            FF.C1=as.numeric(),
                            FF.C2=as.numeric())
i <- 1
for (d in days) {
  temp <- spots.cell.expressing[[1]]
  temp2 <- spots.cell.expressing[[2]]
  temp3 <- spots.cell.expressing[[3]]
  FF.expressing[i,] <- c(var(temp$copy.C0[temp$day==d]) / mean(temp$copy.C0[temp$day==d]),
                                 var(temp2$copy.C1[temp2$day==d]) / mean(temp2$copy.C1[temp2$day==d]),
                                 var(temp3$copy.C2[temp3$day==d]) / mean(temp3$copy.C2[temp3$day==d]))
  i <- i + 1
}



###########################################################################
###########################################################################

Save extracted data to file

write.csv(spots.cell, "spots_cell.csv", row.names=FALSE)
write.csv(spots.cell.2, "spots_cell2.csv", row.names=FALSE)

write.csv(spots.cell.pro, "spots_cell_pro.csv", row.names=FALSE)
write.csv(spots.cell.2.pro, "spots_cell2_pro.csv", row.names=FALSE)

write.csv(TX.molecules, "TX_molecules.csv", row.names=FALSE)


